Non-word repetition in Yélî Dnye

### 2020-11-24 manuscript.Rmd
minor edits: working through intro (incomplete)

- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
<!--M2A: I think you mentioned you might have someone do this check, but let me know if I misunderstood!-->
<!-- a2m we then decided we'd both feel more certain if you did... -->

@@

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood [e.g., @hazan2000development]. One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. NWR scores are thought to reflect long-term phonological knowledge (which allows the child to perceive the item precisely even though it is not a real word they have encountered before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely even though it had never been pronounced before).

NWR has been used to seek answers to a variety of theoretical questions, including what  the links between phonology, working memory, and the lexicon are [@bowey2001nonword], and to assess the extent to which apparent phonological constraints found in the lexicon affect online production [@gallagher2014identity]. It is also frequently used for applications, notably  as a diagnostic for language delays and disorders [@estes2007differences]. Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments,  particularly in Europe [@meir2016disentangling;@is08042009language]. However, it has been seldom used outside of Europe and North America [with exceptions including @gallagher2014identity and @cristia2020infant], and mostly with languages having phonological inventories that are small to moderate in size [@maddieson2005correlating].

<!--M2A: I feel like there's a bit of a jump here from explaining what NWR is and what it's used for to saying we used it in the current study to test three things. We should probably have a paragraph justifying its use in this context, no?-->
<!-- a2m I added a sentence, is this what you were thinking of? -->
<!-- M2A: Makes sense. I was thinking we could already hint at YD as interesting in particular bc of the phonology and not simply because it's a random one of many previously untested populations. I don't mind adding something like this in once I'm done with the demo check.-->
<!-- a2m added ", and mostly with languages having phonological inventories that are small to moderate in size" + notice the " with an unusually dense phonological inventory"; my feeling is that it may interrupt the flow to add a lot more here... don't you think? -->

In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory.  The present report uses these unique data to contribute to several lines of research. First, we  made sure that some of the non-word items contained typologically rare and/or  challenging sounds, so that we can contribute to the study of whether rare sounds are disadvantaged in perception and/or production, both in terms of overall repetition scores and patterns of mispronunciation. Second, we look at the effects of item length, since previous NWR research has uncovered variable effects. Third, we contribute to the basic NWR literature by adding a data point for children learning a rarely-studied language and culture, and further investigating whether there are structured sources that account for individual variation.

>>>>

- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
<!--M2A: Done! Left for AC to review in segments-20201123.xlsx and, if she is satisfied, to integrate :)-->

@@

Children's perception and production of phonetic and phonological units continues developing well beyond the first year of life, even extending into middle childhood [e.g., @hazan2000development]. Much of the evidence for later phonological development comes through the use of nonword repetition (NWR) tasks. In a NWR task, participants hear a short sequence of sounds that is phonologically legal but lexically meaningless in the language(s) they are learning. After hearing this non-word, the participant's task is to try to immediately and precisely repeat it. NWR scores are thought to reflect long-term phonological knowledge (which allows the child to perceive the item precisely despite not having heard it before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely despite not having pronounced it before).

NWR has been used to seek answers to a variety of theoretical questions, including what the links are between phonology, working memory, and the lexicon [@bowey2001nonword], and how extensively apparent phonological constraints found in the lexicon affect online production [@gallagher2014identity]. NWR is also frequently used in applied contexts, notablyas a diagnostic for language delays and disorders [@estes2007differences]. Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe [@meir2016disentangling;@is08042009language]. However, it has seldom been used outside of Europe and North America [with exceptions including @gallagher2014identity and @cristia2020infant], and the languages it has been used with have had small to moderate phonological inventories [@maddieson2005correlating].
<!--M2A: Is the motivating idea here that application to YD is of interest because it's very different from what we've look at with NWR in the past? Or is there a specific hypothesis you had in mind about how inventory size affects phonological development and, hence, NWR performance? I could come up with something, but I think you had an idea already and I don't want to step on it!-->

In the present study, we report on NWR results among children learning Yélî Dnye, an isolate language spoken on Rossel Island, PNG, with a large and unusually dense phonological inventory. The study was designed to contribute to several lines of research: First, we included a subset of non-word items with typologically rare and/or challenging sounds to ask whether these rare sounds are disadvantaged in perception and/or production, both in terms of overall repetition scores and patterns of mispronunciation. Second, we varied the length (in syllables) of non-words to investigate the impact of word length on repetition and mispronunciation, following variable effects of length in previous NWR research. Third, these data give a first impression of how this task is responded to in a rarely-studied language and culture; we add to this by further investigating whether there are structured sources that account for individual variation in our sample.
